Crowd density estimation from Wi-Fi positioning data in the Amsterdam Arena

### 2016-06-22 PaperOnSmoothing/draft-31052016.tex
introduction

Crowd disasters have taken many human lives. The Love Parade disaster (Duisburg, 2010), the Ellis Park Stadium disaster (Johannesburg, 2001), the PhilSports Stadium stampede (Manila, 2006) are just a few recent examples. One of the reasons that disasters happen is a "lack of overview of everybody" [cite paper on love parade EPJ DS], that is, a lack of macroscopic overview of the crowd. Critical crowd density (cite paper 8/m2, find it in e-mails)) is a contributing factor to crowd disasters; yet, it is still challenging to determine timely when it occurs, so that a disaster can be prevented before it happens.

In our work we are interested in estimating crowd density during concerts in indoor spaces. A lot of research on estimating crowd density has been done using video processing from security cameras (cite papers (1), (2), (3) from proposal?). However, this approach does not apply in our case, because 1)  it is difficult to obtaine macroscopic overview ; 2) the lighting conditions during concert hours are not sufficient for video-based crowd analysis; and 3) the error of video-based density estimation increases with the increase of the actual crowd density.

In our approach, we combat the three above mentioned issues with video analysis by exploiting the ubiquity of smart phones, as in ((4) (5), (6) from proposal?). Our living laboratory is the Amsterdam Arena(citatioon). First, we are estimating in real-time anonymously the positions of the visitors via the signal strengths of their phones to the  Wi-Fi routers, as presented in (cite Jan's slides), via a technique similar to trilateration. Second, we apply the law of the large numbers to obtain an estimation of the crowd density, also in real-time. With the first step, we are dealing with issues 1) and 2) mentioned above. With the second step, we are handling issue 3) -  that is, with our method the precision of estimation actually increases as the crowd density increases.

>>>>

Crowd disasters have taken many human lives. The Love Parade disaster (Duisburg, 2010), the Ellis Park Stadium disaster (Johannesburg, 2001), the PhilSports Stadium stampede (Manila, 2006) are just a few recent examples. One of the reasons that disasters happen is a "lack of overview of everybody" [cite paper on love parade EPJ DS], that is, a lack of macroscopic overview of the crowd. Critical crowd density (cite paper 8/m2, find it in e-mails)) is a contributing factor to crowd disasters; yet, it is still challenging to determine timely when it occurs, so that a disaster can be prevented before it happens, by e.g. navigating the rest of the crowd away from the congestion.

In our work we are interested in estimating crowd density during concerts in indoor spaces. A lot of research on estimating crowd density has been done using video processing from security cameras (cite papers (1), (2), (3) from proposal?). However, this approach does not apply in our case, because 1)  it is difficult to obtain macroscopic overview ; 2) the lighting conditions during concert hours are not sufficient for video-based crowd analysis; and 3) the error of video-based density estimation increases with the increase of the actual crowd density.

In our approach, we combat the three above mentioned issues by exploiting the ubiquity of smart phones, as in ((4) (5), (6) from proposal?). Our living laboratory is the Amsterdam Arena(citation). First, we are estimating in real-time anonymously the positions of the visitors via the signal strengths of their phones to the  Wi-Fi routers in the venue. This technique has been presented in (cite Jan's slides), and is similar to trilateration. However, due to signal interference, irregular internet packet rates and systematic errors, this methodology is imprecise as well.  Thus, second, we apply the law of the large numbers to obtain an estimation of the crowd density, also in real-time. With the first step, we are dealing with issues 1) and 2) mentioned above. With the second step, which is the main contribution of this paper, we are handling issue 3) -  that is, with our method the precision of estimation actually increases as the crowd density increases.
